---
title: "class09"
author: "Elisha"
date: "4/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## kmeans clustering
Try the kmeans()

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3)) 
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

```{r}
z <- kmeans(x, centers = 2, nstart=20)
z
```
```{r}
z$size
z$cluster
z$center
```


Use the kmeans() function setting k to 2 and nstart=20 
Inspect/print the results
>Q. How many points are in each cluster?  30/30

```{r}
z$size
```

>Q. What ‘component’ of your result object details - cluster size?
- cluster assignment/membership? - cluster center?

```{r}
z$size
table(z$cluster)
z$center
```


Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
plot(x, col=z$cluster) 
points(z$center, col= "Blue", pch =3, cex= 3)
```


# Hierarchical Clustering
Here we dont have to spell out K the number of clusters before hand but we do have to give it a distance matrix as input.
```{r}
d <-dist(x)
hc <- hclust(d)
hc            
```



Plot the results
```{r}
plot(hc)
abline(h = 6, col = "red")
cutree(hc, h=6)
```

```{r}
gp2 <- cutree(hc, k=2)
gp2
```

```{r}
gp3 <- cutree(hc, k=3)
gp3
```
```{r}
table (gp3)
table(gp2)
table(gp2, gp3)
```

# Try a more real-life like example to see how clusering works
```{r}
x <- rbind(
matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), 
# c1
matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2 
matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3 
         rnorm(50, mean = 0, sd = 0.3)), ncol = 2)) 
colnames(x) <- c("x", "y")

```
```{r}
plot(x)
```
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
>Q Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
dist(x,)
>Q. How does this compare to your known 'col' groups

```{r}
hc1 <- hclust(dist(x))

plot(hc)
```


```{r}
gx2 <-  cutree(hc1, 2)
gx3 <- cutree(hc1,3)

table(gx2)
table(gx3)
table(gx2,gx3)
```


```{r}
plot(x, col=col)

```



##second section : PCA

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names=1)
head(mydata)
```

We will use the base R **prcomp()** function for PCA today...
Get some RNAseq data data to play with
There are `r nrow(mydata)` in this dataset ( will insert the number into the output/knit)
(the upleft key)

```{r}
# t: transpose
head(t(mydata))
# scale will treat each col equally, need the data in row as input, col as variables
pca <- prcomp(t(mydata), scale = T)
summary(pca)
```

```{r}
attributes(pca)
```

Let make out first PCA plot
```{r}
plot(pca$x[,1], pca$x[,2])
```

```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var
```

```{r}
xlab = paste("PC1 (", pca.var.per[1], "%)", sep="")
ylab= paste("PC2 (", pca.var.per[2], "%)", sep="")
plot(pca$x[,1], pca$x[,2], xlab = xlab, ylab=ylab)
```

```{r}
identify(pca$x[,1], pca$x[,2], colnames(mydata))
```


#hands on project
```{r}
x <- read.csv("../UK_foods.csv")
```
Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?
```{r}
nrow(x)
ncol(x)
dim(x)
```

```{r}
head(x, 6)
```

Fixing the row name
```{r}
rownames(x) <- x[,1]
head(x)
x <- x[,-1]
head(x)
```

```{r}
x <- read.csv("../UK_foods.csv", row.names=1)
head(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```
```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(x) )
summary(pca)
```
```{r}
# Plot PC1 vs PC2
mycols <-c("orange", "red", " blue", "darkgreen")
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=mycols)
```


```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
z <- summary(pca)
z$importance
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

