---
title: "Class07"
author: "Elisha"
date: "4/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More on funciton writting
First we will revisit our function from last day.

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
x <- c(1:10, "string")
```


```{r}
rescale2(x)
```

## Function practice
Write a function to identify NA elements in two vectors
Start with a simple example input where I kknow what th eanswer should be
```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c(NA, 3, NA, 3, 4)
```
Looking for the positions where it is TRUE in both vectors
```{r}
sum(is.na(x) & is.na(y))

```
This is the working snippet of code
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```
```{r}
both_na(x,y)
```
```{r}
both_na(c(NA,NA,NA), c(NA,1,NA))
```

```{r}
both_na(c(NA,NA,NA), c(NA,1,NA, NA, NA))
```

Recyclig the short vector......

check the length of our inputs are equal
```{r}
x <- c(NA,NA,NA)
y <- c(NA,1,NA, NA, NA)
length(x)
length(y)

```
Try the both_na3()function with extra feature
```{r}
both_na3(x,y)
```


```{r}
x <- c(100, 100, 100, 100, 100, 100,100, 90)
y <-  c(100, NA, 90, 90,90, 90, 90, 97, 80)
```


```{r}
grade <- function (x) {
  y <- sum(is.na(x)) 
  if (y < 1) {
  (sum(x)-min(x))/(length(x)-1)
  } else{
    (sum(x, na.rm = T) - 0) / (length(x)-1)
  }
}

```
```{r}
z <-  c(100, NA, 90, 90,90, NA, NA, 97, 80)
```
```{r}
grade(x)
grade(y)
grade(z)
```


``{r}
grade2 <- function (x) {
  (sum(x, na.rm =T)-min(x, na.rm =T))/(length(x)-1)
  }


```{r}
url <- "https://tinyurl.com/gradeinput"
students <- read.csv(url, row.names =1)

head(students)
```
```{r}
 grade2(students[5,])
grade2(students[4,])
```

```{r}
ans <- apply(students, 1, grade)
sort(ans, decreasing = T)

```

## Last example
Find the intersection of two sets
```{r}
x <- df1$IDs
y  <- df2$IDs
intersect(x,y)
```

```{r}
x %in% y
which (x %in% y)
x[x %in% y]
```

```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y], y[ y %in% x])
}
```
```{r}
gene_intersect(df1, df2)
```














